|

### 2024-02-19 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
working on rewriting simulations. Will have to rewrite the code for the outcome variable then re-run simulations

To achieve these three goals, the chapter builds upon a growing set of work suggesting that the politicultural differences between Democrats and Republicans are shaping seemingly non-political attitudes and behaviors such as one's choice in which neighborhood to live in. I argue in this chapter that the prevalence of red and blue vehicles in the driveways of houses in a neighborhood shift people's perceptions of the partisan composition of the neighborhood, and as a result, will make people less willing to move to that neighborhood when they perceive they'd be in the political minority of that neighborhood. While these snap-judgments influence attitude expression toward reported willingness to move to a neighborhood, subsequent information such as job opportunities, performance of schools, average wages to cost of living ratio, and other information limits the efficacy of these initial judgments. This seems to provide one mechanism explaining why partisan sorting is not observationally as extreme as we might expect given extensions of the literature on affective polarization and experimental and self-reported survey evidence.

@@

@@

To most of us, we likely do not trust the color of a vehicle as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, political yard signs usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [@makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of vehicle colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red vehicles, we may assume that the local sports team color must be red. However, in line with predictions from Affective Intelligence Theory, states of uncertainty and anxiety predict information seeking behavior [see @marcus_anxiety_1993]. That is, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not. Once we have used that association between the colors red and blue with partisan affiliation, we will form a snap-judgment of our neighbors based on the color of their vehicles when absent other information. That impression will then elicit a positive or negative attitude toward the neighbors and neighborhood as well as encourage a congruent behavioral reaction such as wanting to avoid the neighborhood or an affinity toward it. This impression and behavioral reaction forms pre-consciously and is quickly adjusted with new information from other sources of information. However, the impression constrains what memories are accessible and therefore constrains the attitude expression and behavioral motivations toward objects even when we collect more information that either fits with or goes against our initial impression.

@@

To test these hypotheses, I conduct an experiment involving a fictional real estate posting. The treatment involves randomly assigning participants to view one of two versions of a fictional real estate posting. The characteristics of the neighborhood are exactly the same between the two conditions except for the "favorite color of vehicle" in the neighborhood. In the first condition the "favorite color of vehicle" is reported by displaying a red swatch. In the second condition, the "favorite color of vehicle" reported by displaying a blue swatch. To encourage quick decision making underneath the flyer participants are presented with a feeling thermometer ranging from 0-100 where participants are asked to "Rate their feelings toward the neighborhood as warm (100) or cold (0)". They are also presented with a 10-item scale to indicate, "How much would [they] like to live in this neighborhood?"

@@

Once the participants have responded to these questions, they are then given a "second page" describing the neighborhood which displays as a text box with bullet-points of further information about the neighborhood. The information subjects sees depends on which of two conditions they are randomly assigned into. In the explicitly Republican condition, the information indicates that the neighborhood is "40 minutes from the nearest metro area" and "60% voted for the Republican candidate in the latest congressional election." The second condition acts as a explicitly Democratic condition by presenting the same information but instead describes the neighborhood as "10 minutes from the nearest metro area" and "60% voted for the Democratic candidate in the latest congressional election." With this new information, I ask participants, on the same screen, to "Rate their feelings toward the neighborhood as warm (100) or cold (0)" and are given a feeling thermometer scale to record their response. I also ask them "How much would [they] like to live in this neighborhood?" and are given a 10-item scale to record their response. After answering these questions, I include a manipulation check asking participants what the favorite activity in the neighborhood was with 5 different options: music, movies, football, basketball, and going for drives. Participants who do not answer movies will be excluded from analysis.

@@

The first outcome of interest is the reported feeling thermometer that ranges from 0-100. I divide the response to this outcome variable by 10 and fit an ordered beta regression [see @kubinec_ordered_2022]. As I am interested in examining how these treatments impact perceptions and resulting preferences based on partisanship, I interact the treatment with the self-reported partisan identity of the participant to examine how a participants' preference for that neighborhood is moderated by the treatment. As I am unable to randomly assign my participants' self-reported partisan identity, I want to ensure that I isolate the main and conditional effects of self-reported partisan identification from confounders such as age, and education. I also want to hold constant potential confounders of the treatment's effect on a participant outside of their partisanship such as sex and colorblindness, as well as attention paid to politics. Using these models will allow me to test the first hypothesis which predicts that people will report liking the neighborhood more when the favorite vehicle color aligns with their partisan group -- red with Republicans and blue with Democrats, as they project the association between the colors red and blue with their partisan group onto the neighborhood. I then examine the second hypothesis which predicts that updating this information with consistent cultural stereotypes of partisan preferences for neighborhood features strengthens one's affect towards the neighborhood as a result of this new information strengthening my participants' certainty about the partisan composition of the neighborhood. These tests do not examine whether it influences one's willingness to move to the neighborhood, only one's preferences toward it. In this way, these tests are much like the outcomes examined in the common experimental tests of the observed positive affect toward neighborhoods with a partisan composition aligned with your own identity. The next outcome of interest examines the degree to which someone is willing to move to the neighborhood on this same information.

@@

``` {#lst-sim-dgp-2 .r lst-cap="Code to generate simulated population data (continued)"}
  ...
    RedTreatment = draw_binary( # Simulate treatment assignment with prob 1/3
        N,
        prob = 1/2
    ),
    preference = draw_binary( # Define Preference outcome variable as 3-item ordinal
        x = rnorm(
            N,
            prob = (
              E + 1.5 * RedTreatment + 1.5 * PartyId + 2 * RedTreatment * PartyId + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
            )
        )
    ),
    move = draw_ordered( # Define move outcome variable as Likert variable
        x = rnorm(
            N,
            mean = (
              E + 1 * RedTreatment + 1 * PartyId + 2 * RedTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
            )
        ),
        breaks = c(
            -Inf, -3, -1, 1, 3, Inf
        )
    ),
    preference_post = draw_binary( # Define Preference (post new information) outcome variable as 3-item ordinal
      x = rnorm(
        N,
        prob = (
          E + 1 * RedTreatment + 2 * PartyID + 2.5 * RedTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
        )
      )
    ),
    move_post = draw_ordered(# Define move outcome variable as Likert variable
      x = rnorm(
        N,
        mean = (
          E + 0.5 * RedTreatment + 1.5 * PartyId + 2.5 * RedTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
        )
      ),
      breaks = c(
        -Inf, -3, -1, 1, 3, Inf
      )
    )
)
```

@@

$$
\begin{split}
  \alpha \sim Student-T(3,0,2.5) \\
  \beta_i \sim Normal(0, 1) \\
  \hat{Move_i} = Cumulative(logit^{-1}(\alpha_1 + \alpha_2 + \alpha_3 + \alpha_4 \\ 
   + \beta_1 \times Red Treatment + \beta_2 \times Partisanship \\
   + \beta_3 \times Red treatment \times Partisanship \\ 
   + \beta_4 \times Age + \beta_5 \times Gender \\
   + \beta_6 \times White + \beta_7 \times Attention))
\end{split}
$$ {#eq-party-guess-priors}

@@

if (params$replicate == TRUE) {
  # Define model arguments
  formula <- brmsformula(
    move ~ RedTreatment + PartyID + RedTreatment * PartyID + age + gender + white + attention
  )
  compiled <- stan_model(
    "./STAN/move_model.stan"
    ,model_name = "moveModel"
  )
  family <- cumulative(link = "logit")
  priors <- prior(
    stats::Normal(0,1)
    ,class = b
  )
    # Run model
  modelDiscrepancies <- sim_all(
    n=c(200, 400, 600, 800)
    , num_samples = 500
    , formula = formula
    , compiled = compiled
    , family = family
    , priors = priors
    , cores = cores
    , model = "moveModel"
  )
    # store model
  save(
      modelDiscrepancies
      , file = "../../data/models/prr/sim/sim.RDS"
  )
} else {
  load(
    file = "../../data/models/prr/sim/sim.RDS"
  )
}
```

```{r}
#| label: fig-true-positive-rate
#| layout-nrow: 4
#| fig-width: 6
#| fig-cap: True positive rate
#| fig-subcap: 
#|  - "Red treatment"
#|  - "Party Identification"
#|  - "Red treatment x Party Identification"
#|  - "Age"
#|  - "Gender"
#|  - "White"
#|  - "Attention"
cleanNames <- c(
  "Sample size"
  , "Red treatment"
  , "Partisanship"
  , "Red treatment x Partisanship"
  , "Age"
  , "Gender"
  , "White"
  , "Attention"
)
truePositiveDF <- discrepancy_df(
  list = modelDiscrepancies
  , new_names = cleanNames
)
plotList <- lapply(
  cleanNames
  , true_positive_hist
  , df = truePositiveDF
)
plotList[[2]]
plotList[[3]]
plotList[[4]]
plotList[[5]]
plotList[[6]]
plotList[[7]]
```

>>>>

To achieve these three goals, the chapter builds upon existing work demonstrating our capacity to infer the partisanship of others in low information settings by relying on heuristics built from associations experienced previously [@kam_name_2013], as well as growing set of work suggesting that the politicultural differences between Democrats and Republicans are shaping seemingly non-political attitudes and behaviors -- such as one's choice in which neighborhood to live in. I argue in this chapter that the prevalence of red and blue vehicles in the driveways of houses in a neighborhood shift people's perceptions of the partisan composition of the neighborhood, and as a result, will make people less willing to move to that neighborhood when they perceive they'd be in the political minority of that neighborhood. While these snap-judgments influence attitude expression toward reported willingness to move to a neighborhood, subsequent information such as job opportunities, performance of schools, average wages to cost of living ratio, and other information limits the efficacy of these initial judgments. This seems to provide one mechanism explaining why partisan sorting is not observationally as extreme as we might expect given extensions of the literature on affective polarization and experimental and self-reported survey evidence.

@@

We do not have to actively keep score to inform this behavior. Theories of recognition heuristics suggest that even in low-information settings, like not necessarily knowing the partisan identification of another person, we are capable of inferring the partisanship of another by using associations we have previously made between partisans and their various attributes [@kam_name_2013]. In other words, we do not necessarily need to know that our Uncle Bob identifies as a Democrat, but we can use information about Bob and about the modal Democrat to infer their partisanship.

@@

To most of us, we likely do not trust the color of a vehicle as a basis to infer partisanship as other forms of information such as whether a potential neighbor is displaying a political yard sign. However, political yard signs usually only show up every couple of years during the election season and are displayed by those who tend to be highly politically active -- so, they are not universally displayed [see @kam_name_2013; see also @makse_politics_2019]. Additionally, other decisions such as the color of the house are determined by Homeowners Associations (HOAs). The choices of vehicle colors are less constrained by other external factors. So, we may reasonably assume that if we see a neighborhood with a significant number of red vehicles, we may assume that the local sports team color must be red. However, in line with predictions from Affective Intelligence Theory, states of uncertainty and anxiety predict information seeking behavior [see @marcus_anxiety_1993]. That is, if we are searching for cues toward the partisan composition of the neighborhood, we may then assume that the neighborhood has a large number of Republicans, whether that is accurate or not.

Once we have used that association between the colors red and blue with partisan affiliation, we will form a snap-judgment of our neighbors based on the color of their vehicles when absent other information. That impression will then elicit a positive or negative attitude toward the neighbors and neighborhood as well as encourage a congruent behavioral reaction such as wanting to avoid the neighborhood or an affinity toward it. This impression and behavioral reaction forms pre-consciously and is quickly adjusted with new information from other sources of information. However, the impression constrains what memories are accessible and therefore constrains the attitude expression and behavioral motivations toward objects even when we collect more information that either fits with or goes against our initial impression.

@@

To test these hypotheses, I conduct an experiment involving a fictional real estate posting. I randomly assign half of the participants view the treatment before responding to demographic and political attitude questionnaires containing control variables while the other half view it after. The experiment is a 2x3 factorial design. The treatment involves randomly assigning participants to view one of two versions of a fictional real estate posting. The characteristics of the neighborhood are exactly the same between the two conditions except for the "favorite color of vehicle" in the neighborhood. In the first type of posting the "favorite color of vehicle" is reported by displaying a red swatch. In the second type of posting, the "favorite color of vehicle" reported by displaying a blue swatch. To encourage quick decision making underneath the flyer participants are presented with a feeling thermometer ranging from 0-100 where participants are asked to "Rate their feelings toward the neighborhood as warm (100) or cold (0)". They are also presented with a 10-item scale to indicate, "How much would [they] like to live in this neighborhood?"

@@

Once the participants have responded to these questions, they are then given a "second page" describing the neighborhood which displays as a text box with bullet-points of further information about the neighborhood. There are three sets of information about the neighborhood that participants are randomly assigned to see. In the explicitly Republican condition, the information indicates that the neighborhood is "20 minutes from the nearest metro area" and "60% voted for the Republican candidate in the latest congressional election." The second condition acts as a explicitly Democratic condition by presenting the same information but instead describes the neighborhood as "20 minutes from the nearest metro area" and "60% voted for the Democratic candidate in the latest congressional election." In the final condition, participants are just told that the neighborhood is "20 minutes from the nearest metro area".

With this new information, I ask participants, on the same screen, to "Rate their feelings toward the neighborhood as warm (100) or cold (0)" and are given a feeling thermometer scale to record their response. I also ask them "How much would [they] like to live in this neighborhood?" and are given a 10-item scale to record their response. After answering these questions, I include a manipulation check asking participants what the favorite activity in the neighborhood was with 5 different options: music, movies, football, basketball, and going for drives. Participants who do not answer movies will be excluded from analysis.

@@

The first outcome of interest is the reported feeling thermometer that ranges from 0-100. I divide the response to this outcome variable by 100 and fit an ordered beta regression [see @kubinec_ordered_2022]. As I am interested in examining how these treatments impact perceptions and resulting preferences based on partisanship, I interact the treatment with the self-reported partisan identity of the participant to examine how a participants' preference for that neighborhood is moderated by the treatment. As I am unable to randomly assign my participants' self-reported partisan identity, I want to ensure that I isolate the main and conditional effects of self-reported partisan identification from confounders such as age, and education. I also want to hold constant potential confounders of the treatment's effect on a participant outside of their partisanship such as sex and colorblindness, as well as attention paid to politics. Using these models will allow me to test the first hypothesis which predicts that people will report liking the neighborhood more when the favorite vehicle color aligns with their partisan group -- red with Republicans and blue with Democrats, as they project the association between the colors red and blue with their partisan group onto the neighborhood. I then examine the second hypothesis which predicts that updating this information with consistent cultural stereotypes of partisan preferences for neighborhood features strengthens one's affect towards the neighborhood as a result of this new information strengthening my participants' certainty about the partisan composition of the neighborhood. These tests do not examine whether it influences one's willingness to move to the neighborhood, only one's preferences toward it. In this way, these tests are much like the outcomes examined in the common experimental tests of the observed positive affect toward neighborhoods with a partisan composition aligned with your own identity. The next outcome of interest examines the degree to which someone is willing to move to the neighborhood on this same information.

@@

``` {#lst-sim-dgp-2 .r lst-cap="Code to generate simulated population data (continued)"}
  ...
    RedTreatment = sample( # Simulate treatment assignment with prob 1/3
        1:3,
        N,
        replace=TRUE
    ),
    BlueTreatment = draw_binary( # Simulate treatment assignment with prob 1/3
      N, 
      mean = 1/3
    ),
    preference = draw_binary( # Define Preference outcome variable as 3-item ordinal
        x = rnorm(
            N,
            prob = (
              E + 1.5 * RedTreatment - 1.5 * BlueTreatment + 1.5 * PartyId + 2 * RedTreatment * PartyId - 2 * BlueTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
            )
        )
    ),
    move = draw_ordered( # Define move outcome variable as Likert variable
        x = rnorm(
            N,
            mean = (
              E + 1 * RedTreatment - 1 * BlueTreatment + 1 * PartyId + 2 * RedTreatment * PartyID - 2 * BlueTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
            )
        ),
        breaks = c(
            -Inf, -3, -1, 1, 3, Inf
        )
    ),
    preference_post = draw_binary( # Define Preference (post new information) outcome variable as 3-item ordinal
      x = rnorm(
        N,
        prob = (
          E + 1 * RedTreatment  - 1 * BlueTreatment + 2 * PartyID + 2.5 * RedTreatment * PartyID - 2.5 * BlueTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
        )
      )
    ),
    move_post = draw_ordered(# Define move outcome variable as Likert variable
      x = rnorm(
        N,
        mean = (
          E + 0.5 * RedTreatment - 0.5 * BlueTreatment + 1.5 * PartyId + 2.5 * RedTreatment * PartyID - 2.5 * BlueTreatment * PartyID + 0.5 * age + 0.3 * gender + 0.1 * white + 0.1 * attention
        )
      ),
      breaks = c(
        -Inf, -3, -1, 1, 3, Inf
      )
    )
)
dgp$BlueTreatment <- NA
dgp$BlueTreatment[dgp$RedTreatment == 0] <- 0
dgp$BlueTreatment[dgp$RedTreatment == -1] <- 1
dgp$RedTreatment[dgp$RedTreatment == -1] <- NA
```

@@

$$
\begin{split}
  \alpha \sim Student-T(3,0,2.5) \\
  \beta_i \sim Normal(0, 1) \\
  \hat{Move_i} = Cumulative(logit^{-1}(\alpha_1 + \alpha_2 + \alpha_3 + \alpha_4 \\ 
   + \beta_1 \times Red Treatment + \beta_2 \times Blue Treatment + \beta_3 \times Partisanship \\
   + \beta_4 \times Red treatment \times Partisanship \\ 
   + \beta_5 \times Blue treatment \times Partisanship \\
   + \beta_6 \times Age + \beta_7 \times Gender \\
   + \beta_8 \times White + \beta_9 \times Attention))
\end{split}
$$ {#eq-party-guess-priors}

@@

if (params$replicate == TRUE) {
  # Define model arguments
  formula <- brmsformula(
    move ~ RedTreatment + BlueTreatment + PartyID + RedTreatment * PartyID + BlueTreatment * PartyID + age + gender + white + attention
  )
  compiled <- stan_model(
    "./STAN/move_model.stan"
    ,model_name = "moveModel"
  )
  family <- cumulative(link = "logit")
  priors <- prior(
    stats::Normal(0,1)
    ,class = b
  )
    # Run model
  modelDiscrepancies <- sim_all(
    n=c(200, 400, 600, 800)
    , num_samples = 2
    , formula = formula
    , compiled = compiled
    , family = family
    , priors = priors
    , cores = cores
    , model = "moveModel"
  )
    # store model
  save(
      modelDiscrepancies
      , file = "../../data/models/prr/sim/sim.RDS"
  )
} else {
  load(
    file = "../../data/models/prr/sim/sim.RDS"
  )
}
```

```{r}
#| label: fig-true-positive-rate
#| layout-nrow: 4
#| fig-width: 6
#| fig-cap: True positive rate
#| fig-subcap: 
#|  - "Red treatment"
#|  - "Blue treatment"
#|  - "Party Identification"
#|  - "Red treatment x Party Identification"
#|  - "Blue treatment x Party Identification"
#|  - "Age"
#|  - "Gender"
#|  - "White"
#|  - "Attention"
cleanNames <- c(
  "Sample size"
  , "Red treatment"
  , "Blue treatment"
  , "Partisanship"
  , "Red treatment x Partisanship"
  , "Blue treatment x Partisanship"
  , "Age"
  , "Gender"
  , "White"
  , "Attention"
)
truePositiveDF <- discrepancy_df(
  list = modelDiscrepancies
  , new_names = cleanNames
)
plotList <- lapply(
  cleanNames
  , true_positive_hist
  , df = truePositiveDF
)
plotList[[2]]
plotList[[3]]
plotList[[4]]
plotList[[5]]
plotList[[6]]
plotList[[7]]
plotList[[8]]
plotList[[9]]
```
